Improved Bayesian Ethnorace Prediction

### 2021-05-03 paper.Rmd
final seminar draft

```{r libraries}
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(here)
library(grid)
library(caret)
```

@@

```{=tex}
\begin{singlespace}
Quantitative social science research on race and ethnicity can be constrained by the lack of available individual-level data with these markers. In response to this issue, methods have been developed to predict individuals' race and ethnicity using Bayes' rule. I expand upon existing methods by incorporating new information into the algorithm. Along with some other adjustments, these improvements lead to substantial gains in predictive performance when validated against official state voter files. The largest of these predictive gains are found for African Americans---a group which existing methods find difficult to predict accurately. Furthermore, I apply my algorithm to previous research using the old methods and demonstrate the substantive implications of switching to more accurate predictions.
\end{singlespace}
```
# Introduction

@@

My ethnorace prediction method improves upon Imai and Khanna (2016) in several ways. Whereas their method only takes as inputs distributions over surnames, geolocation, political party, age, and gender, my implementation adds information from a nationwide list of first names [@tzioumis2018] as well as address characteristics. Additionally, I incorporate insights from the machine learning literature to further improve predictive performance. The result of these modifications is a substantial increase in predictive power compared to Imai and Khanna (2016) in validation tests. These predictive gains are particularly strong in regards to correctly classifying African Americans individuals---especially in contexts where fine-grain geolocation data are unavailable. My method is available in an easy-to-use `R` package `bper`.[^1]

@@

In the next section I provide some background and specifics regarding the inputs and outputs of my ethnorace prediction method. Then I explain the methodology and compare my implementation with previous versions. Next, I demonstrate the predictive performance of my method when validated against the combined North Carolina and Florida voter file ($n$ = 21,164,503). And lastly, I replicate Grumbach and Sahn's (2020) descriptive findings in their study on racial disparities in campaign finance using my method of ethnorace prediction. This demonstrates some of the substantive empirical implications of using improved predictions.

@@

There are a few benefits to using the Census ethnorace categorization. These definitions capture a common understanding of race and ethnicity in the US, and correspond to the groups studied most frequently in social science research. The data sources of these groups' distributions that serve as inputs to the prediction formula also rely on the Census categorization. There is currently no feasible alternative data source that I know of that offers the same richness as the Census. This also facilitates comparison of my method against previous ethnorace prediction methods.[^3] One downside to using the Census categories, however, is that it obscures substantial heterogeneity that may exist within each group. Within Asian Americans and Latinos, for example, there is considerable variation in terms of national ancestry. Furthermore, the unfortunate necessity of a catch-all Other Race category ensures that important sources of diversity are lost.[^4] Researchers interested in using this ethnorace prediction method should be aware of these shortcomings. Nonetheless, because the output groups correlate strongly with many important outcomes in social science research, there are still many applications where this prediction method will be useful.

@@

For my last names data, I draw from the 2010 Census Surnames List.[^5] This list comes from the 2010 decennial Census and contains over 160,000 common US last names (those occurring 100 or more times in the population). I also append any last names from the 2000 Census Surnames List that are not found in the 2010 list. This increases the probability the algorithm will match an individual's last name. Like the first names list, these data include counts of individuals in each of the six ethnorace categories across each last name. Once the 2020 Census Surnames List is released, these data will be readily incorporated as well.[^6]

@@

My ethnorace distributions by geographies come from the 2010 decennial Census, accessed via IPUMS NHGIS.[^7] In decreasing order of mean population, these geographies include *state*, *county*, *Census place*, *ZIP code*, and *Census block*. Predictions tend to improve with more precise levels of geography. With this in mind, my implementation automatically matches each individual to the most fine-grain level of geography available. As with the last names input data, once the 2020 Census is released the algorithm will include the updated geolocation distributions.

@@

In general terms, the method computes predicted probabilities for each of the six aforementioned ethnorace categories for each individual. Then, each individual is classified into the category corresponding to the highest predicted probability. These predicted probabilities can be stated more formally as the conditional probability of identifying as a particular ethnorace for an individual with a particular profile of first name, last name, geolocation, party ID, age, gender, and address type. Bayes' rule provides a template for how to answer this sort of conditional probability problem.

@@

Where $R$ is an individual's true ethnorace, $r$ is one of six possible ethnorace categories (White, Black, Asian, Native American, Hispanic, or Other race), and $X$ is the joint probability of an individual having a particular profile of attributes (first name, last name, geolocation, party ID, age, gender, and address type). Unfortunately, the joint probability $X$ in Equation is intractable due to both data constraints and the astronomically large number of combinations of possible attribute profiles. If however, we assume conditional independence of ethnorace among each attribute in $X$, we can rewrite Equation (1) in terms of less complex conditional probabilities:

@@

Fortunately, as I will demonstrate, the predicted probabilities produced by my method are actually very well calibrated. This means that, unlike typical naive Bayes predicted probabilities, they accurately reflect the uncertainty in the algorithm's ethnorace predictions. Therefore researchers should feel confident in incorporating these measures of uncertainty into their analyses when using the method.

@@

To test the performance of the model, I apply the predictions to the combined North Carolina and Florida State voter file.[^12] These files contain snapshots of the registered voters in their respective states and provide individual-level data for first names, last names, address, political party, age, gender, and crucially self-identified ethnorace. After combining the two voter files, I then geocoded each unique address in the sample. This allowed me to match individual observations to Census places and blocks, and ZIP codes. Then I applied the prediction algorithm described above using the `bper` package and calculated each individuals' predicted ethnorace. In order to compare my method against an existing benchmark, I also used the `wru` package [@imai2016] to calculate ethnorace predictions for the same individuals.

@@

Unlike typical machine learning techniques, my prediction method does not fit the model on some subsample, or training set, from these data and then compare predictions against a held-out test set. Instead, the conditional probabilities for each attribute and ethnorace described above are merged into voter file from the input data sources. This allows the entire voter file to be used for validation. And because the input data come from nationally representative samples, the risk of overfitting due to any peculiarities of the North Carolina/Florida electorate are minimized. Together, these two states represent 21,164,503 individuals. Compared to nationwide percentages, Florida has a higher proportion of Hispanics and North Carolina has a higher proportion of African Americans. When combined they form a reasonably ethnoracially diverse population---2% Asian, 16% Black, 11% Hispanic, 6% Other Race, 65% White.

@@

As mentioned previously, the predicted probabilities produced by `bper` are remarkably well calibrated. This means that they accurately reflect the true probability that an individual belongs to a particular ethnorace. We can test this empirically by dividing the predicted probabilities for each ethnorace, for each individual in the sample, into evenly spaced bins (in this case 11). Then we compute the percentage of observed individuals of a particular ethnorace who fall in each bin. If the predicted probabilities are well calibrated, we would expect to see a one-to-one relationship between these two groups. For example, we should observe roughly 25% of individuals whose predicted probability of being White is in the neighborhood of 0.25 to actually be White.

Figure 1 displays the probability calibration plots for Asian, Black, Hispanic, and White predictions in the Florida/North Carolina sample. As demonstrated by their proximity to the dashed 45 degree line, the predictions for Black, Hispanic, and White individuals appear well calibrated. For each bin of predicted probabilities, we observe very close to the corresponding percentage of individuals of that ethnorace in the sample. Calibration for Asian predictions, on the other hand, appear slightly worse. Across the range of predicted probabilities, we observe between 5 to 20% fewer Asian individuals in each bin. This means that the algorithm has a tendency to overestimate Asian predicted probabilities. The comparatively poorer calibration for Asian Americans foreshadows some similar issues with predictive performance for that group in the next section.

```{r, fig.cap="Probability Calibration Plots by Ethnorace"}
cal_api <-
  ggplot(cal_data$bper_cal_api) +
  geom_line(color = "gray20") +
  geom_point(color = "gray20") +
  theme_minimal() +
  labs(y = "Observed Asian Percentage",
       x = "Predicted Probability Asian",
       title = "Asian") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_black <-
  ggplot(cal_data$bper_cal_black) +
  geom_line(color = "gray20") +
  geom_point(color = "gray20") +
  theme_minimal() +
  labs(y = "Observed Black Percentage",
       x = "Predicted Probability Black",
       title = "Black") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_hispanic <-
  ggplot(cal_data$bper_cal_hispanic) +
  geom_line(color = "gray20") +
  geom_point(color = "gray20") +
  theme_minimal() +
  labs(y = "Observed Hispanic Percentage",
       x = "Predicted Probability Hispanic",
       title = "Hispanic") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_white <-
  ggplot(cal_data$bper_cal_white) +
  geom_line(color = "gray20") +
  geom_point(color = "gray20") +
  theme_minimal() +
  labs(y = "Observed White Percentage",
       x = "Predicted Probability White",
       title = "White") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

@@

@@

While investigating the calibration of the predictions is important for assessing the reliability of the predicted probabilities, calibration does not tell us about how well the algorithm ranks these probabilities when generating final ethnorace predictions. Instead, the most straightforward metric for assessing individual-level predictive performance is the Accuracy score, or Overall Error Rate. This number is the proportion of correctly classified individuals in the sample. I ran models separately for different combinations of input variables to mimic data availability constraints in real-world applications, and then calculated the Accuracy score for each. Figure 2 displays a summary of the results of these different models.

@@

ggplot(acc_bper, aes(x = test_type, y = accuracy, group = 1)) +
  geom_line(color = "gray20") +
  geom_point(size = 3, color = "gray20") +
  theme_minimal() +
  ylim(0.75, 0.85) +
  labs(y = "Accuracy Score") +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np")),
                   labels = rev(c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name",
                     "zip" = "ZIP \n Party ID \n First Name \n Last Name",
                     "zip_np" = "ZIP \n First Name \n Last Name",
                     "state" = "State \n Party ID \n First Name \n Last Name",
                     "state_np" = "State \n First Name \n Last Name"
                   ))) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7))
```

As expected, the model with the greatest number of input data sources, and at the most precise geographic level, on the far right of the figure performs the best in terms of overall Accuracy. Using Census blocks, multi-unit address, party ID, first names, and last names, this model correctly identifies the ethnorace of `r round(subset(tests, select = "accuracy", test_method == "bper" & test_type == "all"), 3) * 100`% of individuals in the sample. The upward trend in model accuracy seen in the figure corresponds to shrinking the size of the geolocation variable used. Moving from state to county, from county to place, from place to ZIP code, and from ZIP code to block all improve the overall predictive performance in terms of overall Accuracy. I include the input for multi-unit occupancy only for the Census block models because I believe this reflects the practical contexts where `bper` might be used. If a researcher has access to individual-level addresses, they should be able to geocode these to find the matching Census blocks and should also be able to parse the residency type (multi-unit or stand-alone). But researchers relying on more aggregate geographies likely do not have access to individual addresses, and hence the residency characteristics, for their sample.[^13] In Figure 2 I pair each geolocation variable with a model missing the party ID input. The gains to predictive performance across all geographies appears roughly uniform with the addition of party ID.

@@

Precision and Recall reflect substantively important concerns for real-world applications of the method, and the inherent trade-offs between optimizing for either metric provide a balanced assessment of the predictive performance. On the one hand, Precision rewards very conservative classification procedures. We could, for example, only classify individuals as White if their posterior probability of being White was greater than 99%. This would ensure a very high Precision score for Whites because we are only capturing the low-hanging fruit. A conservative classification procedure like this, however, would likely result in extremely low Recall for Whites. If we only capture the low-hanging fruit, a greater share of White individuals will be mis-classified as non-White. Likewise, optimizing the algorithm for perfect Recall for Whites is trivial. By classifying every individual as White, we ensure 100% of Whites are correctly classified. Of course this procedure would result in extremely low Precision for Whites because every non-White individual would be classified as White as well. Achieving both high Precision and high Recall, therefore, is a difficult task.

@@

ggplot(metrics_bper, aes(x = test_type, y = value, group = Metric)) +
  geom_line(aes(color = Metric), size = 0.3) +
  geom_point(aes(color = Metric), size = 2) +
  theme_minimal() +
  labs(x = "", y = "Score") +
  scale_color_manual(values = c("gray20", "gray50")) +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np")),
                   labels = rev(c(
                     "all" = "Block", 
                     "place" = "Place", 
                     "county" = "County ", 
                     "all_np" = "Block (np)", 
                     "place_np" = "Place (np)", 
                     "county_np" = "County (np)",
                     "zip" = "ZIP",
                     "zip_np" = "ZIP (np)",
                     "state" = "State",
                     "state_np" = "State (np)"
                   ))) +
  facet_wrap(~ Race) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7, angle = 60))
```

Figure 3 displays the Precision and Recall scores broken down by ethnorace for the same models used in Figure 2.[^14] For both Whites and Hispanics, Precision and Recall remain high across all models. Using the most input data available (the models on the far right of the figure), White Precision and Recall are `r round(subset(tests, select = "prec_white", test_method == "bper" & test_type == "all"), 3)` and `r round(subset(tests, select = "rec_white", test_method == "bper" & test_type == "all"), 3)`, respectively. Hispanic Precision and Recall in this model are `r round(subset(tests, select = "prec_hispanic", test_method == "bper" & test_type == "all"), 3)` and `r round(subset(tests, select = "rec_hispanic", test_method == "bper" & test_type == "all"), 3)`, respectively. Hispanic predictive performance across all models likely remains high due to the distinctiveness of Spanish surnames. These metrics are uniformly lower for Asians---likely reflecting the relative rarity of Asian individuals in the sample. When using Census blocks as the unit of geography, predictions for African Americans are quite strong (Precision: `r round(subset(tests, select = "prec_black", test_method == "bper" & test_type == "all"), 3)`, Recall: `r round(subset(tests, select = "rec_black", test_method == "bper" & test_type == "all"), 3)`). But predictive performance falls with broader geolocations. At the state level, without party ID, Black Precision is `r round(subset(tests, select = "prec_black", test_method == "bper" & test_type == "state_np"), 3)` and Recall is `r round(subset(tests, select = "rec_black", test_method == "bper" & test_type == "state_np"), 3)`. This difference in predictive performance between geographies could be explained by the legacy of Black segregation in at least the North Carolina/Florida sample.

@@

```{r}
# metrics_wru <- tests %>% 
#   filter(test_method == "wru") %>% 
#   select(test_type, contains(c("prec", "rec"))) %>% 
#   pivot_longer(cols = contains(c("prec", "rec"))) %>% 
#   separate(name, c("Metric", "Race"), sep = "_") %>% 
#   filter(Race != "other") %>% 
#   mutate(Metric = case_when(Metric == "prec" ~ "Precision",
#                             Metric == "rec" ~ "Recall"),
#          Race = case_when(Race == "api" ~ "Asian",
#                           Race == "black" ~ "Black",
#                           Race == "hispanic" ~ "Hispanic",
#                           Race == "white" ~ "White"))
# 
# ggplot(metrics_wru, aes(x = test_type, y = value, group = Metric)) +
#   geom_line(aes(color = Metric)) +
#   geom_point(aes(color = Metric), size = 1.5) +
#   theme_minimal() +
#   labs(x = "", y = "Score") +
#   scale_color_manual(values = c("gold", "paleturquoise4")) +
#   scale_x_discrete(name = "", limits = c("all", "all_np","place", "place_np", "county", "county_np"),
#                    labels = c(
#                      "all" = "Block", 
#                      "place" = "Place", 
#                      "county" = "County ", 
#                      "all_np" = "Block (np)", 
#                      "place_np" = "Place (np)", 
#                      "county_np" = "County (np)",
#                      "zip" = "ZIP",
#                      "zip_np" = "ZIP (np)",
#                      "state" = "State",
#                      "state_np" = "State (np)"
#                    )) +
#   facet_wrap(~ Race) +
#   theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7, angle = 60))
```

The classification metrics detailed above provide some information about the predictive shortcomings of my method. These limitations notwithstanding, however, the ethnorace predictions made by `bper` are nearly uniformly better than those from `wru`. Figure 4 displays the difference in Accuracy between `bper` and `wru` using the same input data discussed previously.[^15] The baseline of zero in the figure represents the Accuracy scores from each model using `wru`, and the height of the bars display the change in Accuracy using `bper`. Across all model types, `bper` scores roughly between 1.7 and 4.3 percentage points higher on this metric. Regardless of input data, `bper` classifies a higher proportion of individuals correctly in the North Carolina/Florida voter file.

[^15]: The `wru` package does not provide ZIP code or state level predictions so those models are excluded from the comparison.

@@

In "Race and Representation in Campaign Fiance" (2020), Jacob Grumbach and Alexander Sahn investigate racial inequality in campaign contributions. Descriptively, they find that Latinos and African Americans are underrepresented among individuals who donate to US House campaigns. They also use differences-in-differences and regression discontinuity analyses to show that when candidates of color run, the contributor class becomes more ethnoracially representative.

@@

>>>>

```{r libraries}
library(tidyverse)
library(ggthemes)
library(gridExtra)
library(here)
library(grid)
library(caret)
library(wesanderson)
```

@@

```{=tex}
\begin{singlespace}
Quantitative social science research on race and ethnicity can be constrained by the lack of available individual-level data with these markers. In response to this issue, methods have been developed to predict individuals' race and ethnicity using Bayes' rule. I expand upon existing methods by incorporating new information into the algorithm. Along with some other adjustments, these improvements lead to substantial gains in predictive performance when validated against official state voter files. The largest of these predictive gains are found for African Americans---a group which existing methods find difficult to predict accurately. Furthermore, I apply my algorithm to published research using the old methods and demonstrate the substantive implications of switching to more accurate predictions.
\end{singlespace}
```
# Introduction

@@

My ethnorace prediction method improves upon Imai and Khanna (2016) in several ways. Whereas their method only takes as inputs distributions over surnames, geolocation, political party, age, and gender, my implementation adds information from a nationwide list of first names [@tzioumis2018] as well as address characteristics. Additionally, I incorporate insights from the machine learning literature to further improve predictive performance. The result of these modifications is a substantial increase in predictive power compared to Imai and Khanna (2016) in validation tests. These predictive gains are particularly strong in regards to correctly classifying African Americans---especially in contexts where fine-grain geolocation data are unavailable. Under certain data constraints, the probability that an individual predicted to be Black self-identifies as Black is nearly twice as high. My method is available in an easy-to-use `R` package `bper`.[^1]

@@

In the next section I provide some background and specifics regarding the inputs and outputs of my ethnorace prediction method. Then I explain the methodology and compare my implementation with previous versions. Next, I demonstrate the predictive performance of my method when validated against the combined North Carolina and Florida voter file ($n$ = 21,164,503). The inclusion of self-reported ethnorace identifiers in these data allows me to compare my predictions against a ground truth. And lastly, I replicate Grumbach and Sahn's (2020) descriptive findings in their study on racial disparities in campaign finance. I compare the empirical results using my prediction method against those from Imai and Khanna (2016). This demonstrates some of the substantive implications of using improved predictions.

@@

There are several benefits to using the Census ethnorace categorization. These definitions capture a common understanding of race and ethnicity in the US, and correspond to groups studied frequently in social science research. The data sources of these groups' distributions that serve as inputs to the prediction formula also rely on the Census categorization. There is currently no feasible alternative data source that I know of that offers the same level of detail as the Census. This also facilitates comparison of my method against previous ethnorace prediction methods.[^3] One downside to using the Census categories, however, is that it obscures substantial heterogeneity that may exist within each group. Within Asian Americans and Latinos, for example, there is considerable variation in terms of national ancestry. Furthermore, the unfortunate necessity of a catch-all Other Race category, which also includes those who identify as multi-racial, ensures that important sources of diversity are lost.[^4] Researchers interested in using this ethnorace prediction method should be aware of these shortcomings. Nonetheless, because the output groups correlate strongly with many important outcomes in social science research, there are still many applications where this prediction method will be useful.

@@

For my last names data, I draw from the 2010 Census Surnames List.[^5] This list comes from the 2010 decennial Census and contains over 160,000 common US last names (those occurring 100 or more times in the population). I also append any last names from the 2000 Census Surnames List that are not found in the 2010 list to increase the probability the algorithm will match an individual's last name. Like the first names list, these data include counts of individuals in each of the six ethnorace categories across each last name. Once the 2020 Census Surnames List is released, these data will be readily incorporated as well.[^6]

@@

My ethnorace distributions by geolocations come from the 2010 decennial Census, accessed via IPUMS NHGIS.[^7] In decreasing order of mean population, these geographies include *State*, *County*, *Census Place*, *ZIP Code*, and *Census Block*. Predictions tend to improve with more precise levels of geography. With this in mind, my implementation automatically matches each individual to the most fine-grain level of geography available. As with the last names input data, once the 2020 Census data is released the algorithm will include the updated geolocation distributions.

@@

In general terms, the method computes predicted probabilities for each of the six aforementioned ethnorace categories for each individual. Then each individual is classified into the category corresponding to the highest predicted probability. These predicted probabilities can be stated more formally as the conditional probability of identifying as a particular ethnorace for an individual with a particular profile of first name, last name, geolocation, party ID, age, gender, and address type. Bayes' rule provides a template for how to answer this sort of conditional probability problem.

@@

Where $R$ is an individual's true ethnorace, $r$ is one of six possible ethnorace categories (White, Black, Asian, Native American, Hispanic, or Other race), and $X$ is the joint probability of an individual having a particular profile of attributes (first name, last name, geolocation, party ID, age, gender, and address type). Unfortunately, the joint probability $X$ in Equation (1) is intractable due to both data constraints and the astronomically large number of combinations of possible attribute profiles. If however, we assume conditional independence of ethnorace among each attribute in $X$, we can rewrite Equation (1) in terms of less complex conditional probabilities:

@@

Fortunately, as I will demonstrate, the predicted probabilities produced by my method are actually very well calibrated. This means that, unlike typical naive Bayes predicted probabilities which cluster close to 0 and 1 [@zadrozny], the output probabilities from my algorithm accurately reflect the uncertainty in the predictions. Therefore researchers can feel confident in incorporating these measures of uncertainty into their analyses when using the method if they so choose.

@@

To test the performance of the model, I apply the predictions to the combined North Carolina and Florida State voter file.[^12] These files contain snapshots of the registered voters in their respective states and provide individual-level data for first names, last names, address, political party, age, gender, and crucially self-identified ethnorace. After combining the two voter files, I then geocoded each unique address in the sample. This allowed me to match individual observations to Census places and blocks, and ZIP Codes. Then I applied the prediction algorithm described above using the `bper` package and calculated each individuals' predicted ethnorace. In order to compare my method against an existing benchmark, I also used the `wru` package [@imai2016] to calculate ethnorace predictions for the same individuals.

@@

Unlike typical machine learning techniques, my prediction method does not fit the model on some subsample, or training set, from these data and then compare predictions against a held-out test set. Instead, the conditional probabilities for each attribute and ethnorace described in the Inputs section are merged into voter file from the input data sources. This allows the entire voter file to be used for validation. And because the input data come from nationally representative samples, the risk of overfitting due to any peculiarities of the North Carolina/Florida electorate are minimized. Together, these two states represent 21,164,503 individuals. Compared to nationwide percentages, Florida has a higher proportion of Hispanics and North Carolina has a higher proportion of African Americans. When combined they form a reasonably ethnoracially diverse population---2% Asian, 16% Black, 11% Hispanic, 6% Other Race, 65% White.

@@

As mentioned previously, the predicted probabilities produced by `bper` are remarkably well calibrated. This means that they accurately reflect the true probability that an individual belongs to a particular ethnorace. We can test this empirically by dividing the predicted probabilities for each ethnorace, for each individual in the sample, into evenly spaced bins (in this case 11). Then we compute the percentage of observed individuals of a particular ethnorace who fall in each bin. If the predicted probabilities are well calibrated, we would expect to see a one-to-one relationship between these two groups. For example, under good calibration we should observe roughly 25% of individuals whose predicted probability of being White is in the neighborhood of 0.25 to actually be White.

Figure 1 displays the probability calibration plots for Asian, Black, Hispanic, and White predictions in the Florida/North Carolina sample. As shown by their proximity to the dashed 45 degree line, the predictions for Black, Hispanic, and White individuals appear to be well calibrated. For each bin of predicted probabilities, we observe essentially the corresponding percentage of individuals of that ethnorace in the sample. Calibration for Asian predictions, on the other hand, appear somewhat worse. Across the range of predicted probabilities, we observe between 5 to 20% fewer Asian individuals in each bin. This means that the algorithm has a tendency to overestimate Asian predicted probabilities.

```{r, fig.cap="Probability Calibration Plots by Ethnorace"}
cal_api <-
  ggplot(cal_data$bper_cal_api) +
  geom_line(color = wes_palette("Moonrise2", n = 1)) +
  geom_point(color = wes_palette("Moonrise2", n = 1)) +
  theme_minimal() +
  labs(y = "Observed Asian Percentage",
       x = "Predicted Probability Asian",
       title = "Asian") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_black <-
  ggplot(cal_data$bper_cal_black) +
  geom_line(color = wes_palette("Moonrise2", n = 1)) +
  geom_point(color = wes_palette("Moonrise2", n = 1)) +
  theme_minimal() +
  labs(y = "Observed Black Percentage",
       x = "Predicted Probability Black",
       title = "Black") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_hispanic <-
  ggplot(cal_data$bper_cal_hispanic) +
  geom_line(color = wes_palette("Moonrise2", n = 1)) +
  geom_point(color = wes_palette("Moonrise2", n = 1)) +
  theme_minimal() +
  labs(y = "Observed Hispanic Percentage",
       x = "Predicted Probability Hispanic",
       title = "Hispanic") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

cal_white <-
  ggplot(cal_data$bper_cal_white) +
  geom_line(color = wes_palette("Moonrise2", n = 1)) +
  geom_point(color = wes_palette("Moonrise2", n = 1)) +
  theme_minimal() +
  labs(y = "Observed White Percentage",
       x = "Predicted Probability White",
       title = "White") +
  ylim(0, 100) +
  theme(text = element_text(family = "serif"), plot.title = element_text(hjust = 0.5))

@@

Given the discovery of over-inflated predicted probabilities for Asian Americans, I subsequently adjusted the algorithm to down-weight these values by 50%. Reducing the initial Asian predicted probabilities *greatly* increased overall predictive performance for this group without influencing the results for other ethnoraces. While the down-weighting for Asian predictions can be defended on the basis of the calibration plots in Figure 1, even better predictive results could be obtained through a more structured approach to adjusting predicted probabilities post-hoc. Platt scaling and isotonic regression are two methods developed to improve the calibration, and the subsequent predictive performance, of similar algorithms [@zadrozny].

@@

While investigating the calibration of the predictions is important for assessing the reliability of the predicted probabilities, calibration does not tell us how well the method ranks these probabilities when generating final ethnorace predictions. Instead, the a common metric for assessing individual-level predictive performance is the Accuracy score, or Overall Error Rate. This number is the proportion of correctly classified individuals in the sample. I ran models separately for different combinations of input variables to mimic data availability constraints in real-world applications, and then calculated the Accuracy score for each. Figure 2 displays a summary of the results of these different models.

@@

ggplot(acc_bper, aes(x = test_type, y = accuracy, group = 1)) +
  geom_line(color = wes_palette("Moonrise2", n = 1)) +
  geom_point(size = 3, color = wes_palette("Moonrise2", n = 1)) +
  theme_minimal() +
  ylim(0.75, 0.85) +
  labs(y = "Accuracy Score") +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np")),
                   labels = rev(c(
                     "all" = "Block \n Multi-Unit \n Party ID \n First Name \n Last Name", 
                     "place" = "Place \n Party ID \n First Name \n Last Name", 
                     "county" = "County \n Party ID \n First Name \n Last Name", 
                     "all_np" = "Block \n Multi-Unit \n First Name \n Last Name", 
                     "place_np" = "Place \n First Name \n Last Name", 
                     "county_np" = "County \n First Name \n Last Name",
                     "zip" = "ZIP \n Party ID \n First Name \n Last Name",
                     "zip_np" = "ZIP \n First Name \n Last Name",
                     "state" = "State \n Party ID \n First Name \n Last Name",
                     "state_np" = "State \n First Name \n Last Name"
                   ))) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7))
```

As expected, the model with the greatest number of input data sources, and at the most precise geographic level, on the far right of the figure performs the best in terms of overall Accuracy. Using Census blocks, multi-unit address, party ID, first names, and last names, this model correctly identifies the ethnorace of `r round(subset(tests, select = "accuracy", test_method == "bper" & test_type == "all"), 3) * 100`% of individuals in the sample. The upward trend in model accuracy seen in the figure corresponds to shrinking the size of the geolocation variable used. Moving from state to county, from county to place, from place to ZIP code, and from ZIP code to block all improve the overall predictive performance in terms of overall Accuracy. I include the input for multi-unit occupancy only for the Census block models because this reflects the practical contexts where `bper` might be used. If a researcher has access to individual-level addresses, they should be able to geocode these to find the matching Census blocks and should also be able to parse the residency type (multi-unit or stand-alone). But researchers relying on more aggregate geographies likely do not have access to individual addresses, and hence the residency characteristics, for their sample.[^13] In Figure 2 I pair each geolocation variable with a model missing the party ID input. The predictive gains from using party ID appear to diminish slightly with smaller levels of geography.

@@

Precision and Recall reflect substantively important concerns for real-world applications of the method, and the inherent trade-offs between optimizing for either metric provide a balanced assessment of the predictive performance. On the one hand, Precision rewards very conservative classification procedures. We could, for example, only classify individuals as White if their predicted probability of being White was greater than 99%. This would ensure a very high Precision score for Whites because we are only capturing the low-hanging fruit. A conservative classification procedure like this, however, would likely result in extremely low Recall for Whites. If we only capture the low-hanging fruit, a greater share of White individuals will be mis-classified as non-White. Likewise, optimizing the algorithm for perfect Recall for Whites is trivial. By classifying every individual as White, we ensure 100% of Whites are correctly classified. Of course this procedure would result in extremely low Precision for Whites because every non-White individual would be classified as White as well. Achieving both high Precision and high Recall, therefore, is a difficult task.

@@

ggplot(metrics_bper, aes(x = test_type, y = value, group = Metric)) +
  geom_line(aes(color = Metric), size = 0.3) +
  geom_point(aes(color = Metric), size = 2) +
  theme_minimal() +
  labs(x = "", y = "Score") +
  scale_color_manual(values = wes_palette("Moonrise2", n = 3)) +
  scale_x_discrete(name = "Model Inputs", limits = rev(c("all", "all_np", "zip", "zip_np", "place", "place_np", "county", "county_np", "state", "state_np")),
                   labels = rev(c(
                     "all" = "Block", 
                     "place" = "Place", 
                     "county" = "County ", 
                     "all_np" = "Block (np)", 
                     "place_np" = "Place (np)", 
                     "county_np" = "County (np)",
                     "zip" = "ZIP",
                     "zip_np" = "ZIP (np)",
                     "state" = "State",
                     "state_np" = "State (np)"
                   ))) +
  facet_wrap(~ Race) +
  theme(text = element_text(family = "serif"), axis.text.x = element_text(size = 7, angle = 60))
```

Figure 3 displays the Precision and Recall scores broken down by ethnorace for the same models used in Figure 2.[^14] For both Whites and Hispanics, Precision and Recall remain high across all models. Using the most input data available (the models on the far right of the figure), White Precision and Recall are `r round(subset(tests, select = "prec_white", test_method == "bper" & test_type == "all"), 3)` and `r round(subset(tests, select = "rec_white", test_method == "bper" & test_type == "all"), 3)`, respectively. Hispanic Precision and Recall in this model are `r round(subset(tests, select = "prec_hispanic", test_method == "bper" & test_type == "all"), 3)` and `r round(subset(tests, select = "rec_hispanic", test_method == "bper" & test_type == "all"), 3)`, respectively. Hispanic predictive performance across all models likely remains high due to the distinctiveness of Spanish surnames. These metrics are uniformly lower for Asians---likely reflecting the relative rarity of Asian individuals in the sample and the heterogeneous qualities of this particular group.[^15] When using Census blocks as the unit of geography, predictions for African Americans are quite strong (Precision: `r round(subset(tests, select = "prec_black", test_method == "bper" & test_type == "all"), 3)`, Recall: `r round(subset(tests, select = "rec_black", test_method == "bper" & test_type == "all"), 3)`). But predictive performance falls with broader geolocations. At the state level, without party ID, Black Precision is `r round(subset(tests, select = "prec_black", test_method == "bper" & test_type == "state_np"), 3)` and Recall is `r round(subset(tests, select = "rec_black", test_method == "bper" & test_type == "state_np"), 3)`. This difference in predictive performance between geographies could be explained by the legacy of Black segregation in at least the North Carolina/Florida sample.

@@

[^15]: I investigated which ethnoraces Asian Americans were wrongly predicted to be and found a wide spread: 65% White, 18% Hispanic, 14% Black, 1% Other

The classification metrics detailed above provide some information about the predictive shortcomings of my method. These limitations notwithstanding, however, the ethnorace predictions made by `bper` are nearly uniformly better than those from `wru` [@imai2016]. Figure 4 displays the difference in Accuracy between `bper` and `wru` using the same input data discussed previously.[^16] The baseline of zero in the figure represents the Accuracy scores from each model using `wru`, and the height of the bars display the change in Accuracy using `bper`. Across all model types, `bper` scores between 1.8 and 4.6 percentage points higher on this metric. Regardless of input data, `bper` classifies a higher proportion of individuals correctly in the North Carolina/Florida voter file.

[^16]: The `wru` package does not provide ZIP code or state level predictions so those models are excluded from the comparison.

@@

In "Race and Representation in Campaign Fiance" (2020), Jacob Grumbach and Alexander Sahn investigate racial inequality in campaign contributions. Descriptively, they find that Latinos and African Americans are underrepresented among individuals who donate to US House campaigns. They also use differences-in-differences and regression discontinuity analyses to show that when candidates of color run, the contributor class becomes more ethnoracially representative because these candidates garner more co-ethnic donations.

@@

This paper describes a powerful method for predicting individuals' race or ethnicity based on other known attributes. Through validation tests, I have shown that my method produces better predictions than those used in recent social science research. These improvements are not trivial. If researchers only have access to individuals' name and county or place, my method nearly doubles the probability that someone predicted to be Black self-identifies as Black relative to predictions generated from other methods [@imai2016]. These improved predictions reveal new empirical findings as demonstrated by my replication of Grumbach and Sahn (2020).
